Response Patterns

### 2024-11-04 misc/discrep.qmd
update with before and after discrepancy exclusion tables and figures

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,tibble, haven,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

@@

>>>>

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,tibble, haven,
               ggplot2,gt,knitr,kableExtra,ggh4x,ggdist,gghalves,patchwork,lme4,flextable)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

@@

## Data after exclusions

### Study 1

# number of unique ids per condition
s1 |> group_by(refClass) |> 
filter(!(id %in% s1_discrep_ids)) |>
rename("Reference Class" = refClass) |>
summarise(n=n_distinct(id)) |> gt() |> tab_header(title="Study 1: Number of unique IDs per condition")

s1 |> group_by(refClass,state) |> 
  filter(!(id %in% s1_discrep_ids)) |>
  rename("Reference Class" = refClass) |>
  summarise(n=n_distinct(id)) |> 
  rowwise() |> 
  mutate(state=lookup_code(state)) |> 
  pivot_wider(names_from=state,values_from=n) |>
  gt() |> tab_header(title="Study 1: Number of unique IDs per condition by state")

s1 |> group_by(refClass,calc) |> 
  filter(!(id %in% s1_discrep_ids)) |>
  rename("Reference Class" = refClass) |>
  summarise(n=n_distinct(id)) |> 
  pivot_wider(names_from=calc,values_from=n) |>
  gt() |> tab_header(title="Study 1: Number of unique IDs per condition by calculation")

s2_long |> group_by(refClass) |>
filter(!(id %in% s2_discrep_ids)) |>
rename("Reference Class" = refClass) |>
summarise(n=n_distinct(id)) |> gt() |> tab_header(title="Study 2: Number of unique IDs per condition")

s2_long |> group_by(refClass,calc) |>
filter(!(id %in% s2_discrep_ids)) |>
rename("Reference Class" = refClass) |>
mutate(calc=case_when(calc=="Calculator" ~ "Used Calculator",
                      calc=="No Calculator" ~ "No Calculator")) |>
summarise(n=n_distinct(id)) |>
pivot_wider(names_from=calc,values_from=n) |>
gt() |> tab_header(title="Study 2: Number of unique IDs per condition by calculation")

# s2_long |> group_by(id) |> summarise(nrc=n_distinct(refClass),npct=n_distinct(pct_goal),nround=n_distinct(rounded))

### Performance before and after exclusions - Study 1

```{r}
#| fig-width: 12
#| fig-height: 8

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
            n_change = sum(value!=family),
            state_p_dif=mean(state_p_dif),
            state_f_dif=mean(state_f_dif),
            n_less_avg = sum(less_avg),
            duration=first(Duration__in_seconds_)) |> 
  mutate(matched_goal = (pct_change == pct_goal), 
             error = pct_change - pct_goal,
            abs_error = abs(error),
            close_match = abs_error <= 0.02)

s1_agg4 <- s1_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

# s1_agg4 |> group_by(refClass) |> 
#   summarise(mean=mean(pct),sd=sd(pct),n=n()) |> 
#   gt() |> tab_header(title="Study 1: Performance before exclusions")

# s1_agg4 |> 
#   filter(!(id %in% s1_discrep_ids)) |> 
#   group_by(refClass) |> 
#   summarise(mean=mean(pct),sd=sd(pct),n=n()) |> 
#   gt() |> tab_header(title="Study 1: Performance after exclusions")

bind_rows(
  s1_agg4 |> 
    group_by(refClass) |> 
    summarise(
      mean = mean(pct),
      sd = sd(pct),
      n = n(),
      dataset = "Before exclusions"
    ),
  s1_agg4 |> 
    filter(!(id %in% s1_discrep_ids)) |> 
    group_by(refClass) |> 
    summarise(
      mean = mean(pct),
      sd = sd(pct),
      n = n(),
      dataset = "After exclusions"
    )
) |>
  pivot_wider(
    names_from = dataset,
    values_from = c(mean, sd, n),
    names_glue = "{dataset}_{.value}"
  ) |>
  gt() |>
  tab_header(
    title = "Study 1: Performance Before and After Exclusions"
  ) |>
  fmt_number(
    columns = contains("mean"),
    decimals = 2
  ) |>
  fmt_number(
    columns = contains("sd"),
    decimals = 2
  ) |>
  cols_label(
    refClass = "Reference Class",
    `Before exclusions_mean` = "Mean",
    `Before exclusions_sd` = "SD",
    `Before exclusions_n` = "N",
    `After exclusions_mean` = "Mean",
    `After exclusions_sd` = "SD",
    `After exclusions_n` = "N"
  ) |>
  tab_spanner(
    label = "Before Exclusions",
    columns = starts_with("Before")
  ) |>
  tab_spanner(
    label = "After Exclusions",
    columns = starts_with("After")
  )

sp1b <- s1_agg4 |> 
  ggplot(aes(x=refClass,y=pct,fill=refClass)) +
  stat_summary(fun=mean, geom="bar") +
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
  labs(title="Mean % meeting goal - Before Exclusions",
       x="Reference Class",
       y="% of trials meeting goal",
       fill="Reference Class")

sp1a <- s1_agg4 |> 
  filter(!(id %in% s1_discrep_ids)) |>
  ggplot(aes(x=refClass,y=pct,fill=refClass)) +
  stat_summary(fun=mean, geom="bar") +
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
    labs(title="Mean % meeting goal - After Exclusions",
       x="Reference Class",
       y="% of trials meeting goal",
       fill="Reference Class")

sp2b <- s1_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=pct)) +
  geom_bar(aes(fill=refClass)) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  labs(title="Distributions of % of trials that matched goal - Before Exclusions",
       x="% of trials (out of 4) that matched goal",
       y="Number of Participants",
       fill="Reference Class") +
  scale_y_continuous(limits=c(0,85),breaks=seq(0,85,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))

sp2a <- s1_agg |> 
  filter(!(id %in% s1_discrep_ids)) |> 
  group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=pct)) +
  geom_bar(aes(fill=refClass)) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  labs(title="Distributions of % of trials that matched goal - After Exclusions",
       x="% of trials (out of 4) that matched goal",
       y="Number of Participants",
       fill="Reference Class") +
  scale_y_continuous(limits=c(0,85),breaks=seq(0,85,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))

sp1b + sp1a

sp2b + sp2a

### After Exclusions - Study 2

```{r}
#| fig-width: 12
#| fig-height: 8

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, refClass,pct,plan) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
  matched_goal2 = abs(pct_change-pct)<.03)

s2_agg4 <- s2_agg1|> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

bind_rows(
  s2_agg4 |> 
    group_by(refClass) |> 
    summarise(
      mean = mean(pct),
      sd = sd(pct),
      n = n(),
      dataset = "Before exclusions"
    ),
  s2_agg4 |> 
    filter(!(id %in% s2_discrep_ids)) |> 
    group_by(refClass) |> 
    summarise(
      mean = mean(pct),
      sd = sd(pct),
      n = n(),
      dataset = "After exclusions"
    )) |> 
    pivot_wider(names_from = dataset,
    values_from = c(mean, sd, n),
    names_glue = "{dataset}_{.value}") |> gt() |>
    tab_header(title = "Study 2: Performance Before and After Exclusions") |>
    cols_label(
    refClass = "Reference Class",
    `Before exclusions_mean` = "Mean",
    `Before exclusions_sd` = "SD",
    `Before exclusions_n` = "N",
    `After exclusions_mean` = "Mean",
    `After exclusions_sd` = "SD",
    `After exclusions_n` = "N"
  ) |>
  tab_spanner(
    label = "Before Exclusions",
    columns = starts_with("Before")
  ) |>
  tab_spanner(
    label = "After Exclusions",
    columns = starts_with("After"))

s2p1b <- s2_agg4 |> 
  ggplot(aes(x=refClass,y=pct,fill=refClass)) +
  stat_summary(fun=mean, geom="bar") +
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
  labs(title="Mean % meeting goal - Before Exclusions",
       x="Reference Class",
       y="% of trials meeting goal",
       fill="Reference Class")

s2p1a <- s2_agg4 |> 
  filter(!(id %in% s2_discrep_ids)) |>
  ggplot(aes(x=refClass,y=pct,fill=refClass)) +
  stat_summary(fun=mean, geom="bar") +
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
  labs(title="Mean % meeting goal - After Exclusions",
       x="Reference Class",
       y="% of trials meeting goal",
       fill="Reference Class")

s2p2b <- s2_agg1 |> group_by(id,refClass) |>
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=pct)) +
  geom_bar(aes(fill=refClass)) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  labs(title="Distributions of % of trials that matched goal - Before Exclusions",
       x="% of trials (out of 4) that matched goal",
       y="Number of Participants",
       fill="Reference Class") + 
       # make title text smaller
        theme(plot.title = element_text(size=10)) +
  scale_y_continuous(limits=c(0,85),breaks=seq(0,85,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))

s2p2a <- s2_agg1 |> 
  filter(!(id %in% s2_discrep_ids)) |> 
  group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
  ggplot(aes(x=pct)) +
  geom_bar(aes(fill=refClass)) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  facet_wrap(~refClass) +
  labs(title="Distributions of % of trials that matched goal - After Exclusions",
       x="% of trials (out of 4) that matched goal",
       y="Number of Participants",
       fill="Reference Class") +
  scale_y_continuous(limits=c(0,85),breaks=seq(0,85,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))

s2p1b + s2p1a

s2p2b + s2p2a

### 2024-11-04 misc/indiv.qmd
update with before and after discrepancy exclusion tables and figures

## Study 1 - comparison against state average
```{r}

s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
  filter(appliance!="TOTAL",state=="California") |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

>>>>

## Study 1 - comparison against state average

```{r}
s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
  filter(appliance!="TOTAL",state=="California") |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
